An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects in Time-Frequency Representations

### 2020-07-14 Notes/results.tex
code

\begin{table}
\centering
\caption{AHM signal. Performance of the extension methods.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
   \cline{2-3} & Mean & Standard deviation & \\
   \hhline{|=#=|=|=|}
   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.152$ \\
   \hline
   Symmetric & $1.019\times 10^{1}$ & $1.192\times 10^{2}$ & $0.002$ \\
   \hline
   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.537$\\
   \hline
   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.331$ \\
   \hline
   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.120$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

@@

In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given TF representation and the optimal one. It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions. Let us generically denote a time frequency representation $\ccQ$. Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left. |\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$. At each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two densities. It is given by the $L^1$ norm of the difference between the associated distribution functions. In other words, we have:
\begin{equation*}
d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ ,
\end{equation*}
where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$. The \textit{optimal transport distance} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation. That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the averaged ratio between the optimal transport distance to the optimal TF representation~\eqref{eq:opt.TFR} and the optimal transport distance of to the optimal TF representation, that is:
\begin{equation}
D(\ccQ) = \dfrac1{|I|}\bigintss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
\label{eq:index.perf}
\end{equation}
Thus, $D(\ccQ)<1$ means a reduction of the boundary effects. Let us evaluate the quality of the boundary effects reduction on biomedical signals.

@@

\begin{table}
\centering
\caption{Performance of the extension methods on a respiratory signal.}
\begin{tabular}{|c||c||c|c|c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multirow{2}{35pt}{\centering Averaged MSE} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
   \cline{3-6}
      & & STFT & SST & RS & ConceFT \\
   \hhline{|=#=#=|=|=|=|}
   {\sf SigExt} & $0.046$ & $0.705$ & $0.698$ & $0.694$ & $0.776$ \\
   \hline
   Symmetric & $0.039$ & $1.280$ & $0.984$ & $9.063$ & $1.295$ \\
   \hline
   EDMD & $0.022$ & $0.757$ & $0.635$ & $0.701$ & $0.793$ \\
   \hline
   GPR & $0.045$ & $0.817$ & $0.793$ & $0.727$ & $0.876$ \\ 
   \hline
\end{tabular}
\label{tab:THO}
\end{table}

@@

\begin{table}
\centering
\caption{PPG signal: averaged performance index of the boundary-free TF representations for different extensions.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Performance index $D$} \\
   \cline{2-4}
      & STFT & SST & ConceFT\\
   \hhline{|=#=|=|=|}
%   Without extension & $2.52\times 10^{-2}$ & $9.41\times 10^{-2}$ & $1.03\times 10^{-1}$ \\
%   \hline
   {\sf SigExt} & $0.774$ & $0.048$ & $0.826$ \\
   \hline
   Symmetric & $0.818$ & $0.059$ & $1.393$ \\
   \hline
   EDMD & $0.566$ & $0.038$ & $0.850$ \\
   \hline
   GPR & $0.735$ & $0.045$ & $0.816$ \\
   \hline
\end{tabular}
\label{tab:otd.ppg}
\end{table}

>>>>

\begin{table}
\centering
\caption{AHM signal: performance of the extension methods.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
   \cline{2-3} & Mean & Standard deviation & \\
   \hhline{|=#=|=|=|}
   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.152$ \\
   \hline
   Symmetric & $1.019\times 10^{1}$ & $1.192\times 10^{2}$ & $0.002$ \\
   \hline
   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.537$\\
   \hline
   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.331$ \\
   \hline
   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.120$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

@@

In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given TF representation and the optimal one. It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions. Let us generically denote a time frequency representation $\ccQ$. Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left. |\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$. At each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two densities. It is given by the $L^1$ norm of the difference between the associated distribution functions. In other words, we have:
\begin{equation*}
d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ ,
\end{equation*}
where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$. The \textit{Optimal Transport Distance (OTD)} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation. That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the ratio between its averaged OTD to the optimal TF representation $\ccF^\mathrm{opt}$ (see~\eqref{eq:opt.TFR}) and the averaged OTD of the original TF representation $\ccF$ to $\ccF^\mathrm{opt}$, that is:
%\begin{equation}
%D(\ccQ) = \dfrac1{|I|}\bigintssss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
%\label{eq:index.perf}
%\end{equation}
\begin{equation}
D(\ccQ) = \dfrac{\displaystyle\int_I d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)\dd t}{\displaystyle\int_I d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)\dd t}\ .
\label{eq:index.perf}
\end{equation}
Thus, $D(\ccQ)<1$ means a reduction of the boundary effects. Let us evaluate the quality of the boundary effects reduction on biomedical signals.

@@

\begin{table}
\centering
\caption{Respiratory signal: Performance of extension methods and associated boundary-free TF representations.}
\begin{tabular}{|c||c||c|c|c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{35pt}{\centering Averaged MSE} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
   \cline{3-6}
      & & STFT & SST & RS & ConceFT \\
   \hhline{|=#=#=|=|=|=|}
   {\sf SigExt} & $0.046$ & $0.705$ & $0.698$ & $0.694$ & $0.776$ \\
   \hline
   Symmetric & $0.039$ & $1.280$ & $0.984$ & $9.063$ & $1.295$ \\
   \hline
   EDMD & $0.022$ & $0.757$ & $0.635$ & $0.701$ & $0.793$ \\
   \hline
   GPR & $0.045$ & $0.817$ & $0.793$ & $0.727$ & $0.876$ \\ 
   \hline
\end{tabular}
\label{tab:THO}
\end{table}

@@

\begin{table}
\centering
\caption{PPG signal: Performance of the boundary-free TF representations according to the extension method.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{3}{c|}{Averaged performance index $D$} \\
   \cline{2-4}
      & STFT & SST & ConceFT\\
   \hhline{|=#=|=|=|}
%   Without extension & $2.52\times 10^{-2}$ & $9.41\times 10^{-2}$ & $1.03\times 10^{-1}$ \\
%   \hline
   {\sf SigExt} & $0.774$ & $0.048$ & $0.826$ \\
   \hline
   Symmetric & $0.818$ & $0.059$ & $1.393$ \\
   \hline
   EDMD & $0.566$ & $0.038$ & $0.850$ \\
   \hline
   GPR & $0.735$ & $0.045$ & $0.816$ \\
   \hline
\end{tabular}
\label{tab:otd.ppg}
\end{table}
